---
title: "ANOVA_Multiple Comparisons"
author: "Conor Chandler"
output: 
    md_document:
      variant: markdown_github
      toc: true
editor_options:
  chunk_output_type: console
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r}
#Need for Dunnett distribution
library(nCDunnett)
```

Stacked format of data from Lesson 1 was saved tor a csv file and read into an R dataframe
```{r}
df <- read.csv("StackedData.csv",header = F)
```

Define variables for use in subsequent steps
```{r}
#Four treatments
NumOfGroups <- 4
#Number of pairwise comparisons is simply 4 choose 2
NumOfCompar <- factorial(NumOfGroups) / factorial(NumOfGroups-2) / factorial(2)
#Number of observations per treatment is 6
NumofRepl <- 6
#Set alpha to be 5%
alpha <- 0.05
```

Fit an ANOVA model and use R's TukeyHSD function to carryout the multiple comparison procedure
```{r}
aov_Fert <- aov(V2 ~ V1, data = df)
(tuk_Res <- TukeyHSD(aov_Fert))
```

Extract the mean differences from tuk_Res and create a new data frame called pairs_df
```{r}
pairs_df <- as.data.frame(sapply(tuk_Res, function(x) x[,1]))
```

Extract MSE and error DF from ANOVA table, and compute the standard error for the difference between two treatment means.
From Kutner et al. (17.30b), the standard error for pairwise comparison $D=\mu_i-\mu_i'$ is $s[\hat{D}]=\sqrt{MSE(1/n_i+1/n_i')}$. The number of observations is equal in each group, so this could be simplified to $s[\hat{D}]=\sqrt{2MSE/n}$.
```{r}
summary(aov_Fert)
myMSE <- summary(aov_Fert)[1][[1]][[3]][[2]]
errorDF <- summary(aov_Fert)[1][[1]][[1]][[2]]
se_pair <- sqrt(myMSE*((1/NumofRepl)+(1/NumofRepl)))
```

# Comparison of Multiple Comparison Methods
All of them produce simultaneous confidence intervals of the general form: sample statistic +/- margin of error, where margin of error = multiplier * SE(estimate). All of the methods differ in the form of the multiplier

## Fisher’s Least Significant Difference (LSD)
```{r}
LSD_crit <- qt(1-alpha/2, df = errorDF)
MofE <- se_pair*LSD_crit
pairs_df$lowerLSD <- pairs_df$V1-MofE
pairs_df$upperLSD <- pairs_df$V1+MofE
```

## Tukey
```{r}
qtukey_val <- qtukey(1-alpha, NumOfGroups, df = errorDF)/sqrt(2)
MofE <- se_pair*qtukey_val
pairs_df$lowerTuk <- pairs_df$V1-MofE
pairs_df$upperTuk <- pairs_df$V1+MofE
```

P-values match the TukeyHSD output (right-tail probabilities)
Note, the test statistic is $q^*=\sqrt{2}\hat{D}/s[\hat{D}]$ (Kunter et al., 17.32)
```{r}
ptukey(abs(pairs_df$V1*sqrt(2)/se_pair), NumOfGroups, df=errorDF,lower.tail = F)
```

## Bonferroni
```{r}
Bon_crit <- qt(1-alpha/2/NumOfCompar, df = errorDF)
MofE <- se_pair*Bon_crit
pairs_df$lowerBon <- pairs_df$V1-MofE
pairs_df$upperBon <- pairs_df$V1+MofE
```

## Scheffe
```{r}
S_crit <- sqrt((NumOfGroups-1)*qf(1-alpha, df1=NumOfGroups-1, df2=errorDF))
MofE <- se_pair*S_crit
pairs_df$lowerS <- pairs_df$V1-MofE
pairs_df$upperS <- pairs_df$V1+MofE
```

## Dunnett
```{r}
Dun_crit <- qNCDun(p=1-alpha, nu=errorDF, rho=(rep(0.5,times=NumOfGroups-1)), delta=rep(0,times=NumOfGroups-1), two.sided=T)
MofE <- se_pair*Dun_crit
pairs_df$lowerDun <- pairs_df$V1-MofE
pairs_df$upperDun <- pairs_df$V1+MofE
pairs_df$lowerDun[4:6] <- NA
pairs_df$upperDun[4:6] <- NA
```

Comparison of the 95% simultaneous confidence intervals produced
```{r}
knitr::kable(pairs_df, digits = 3)
```

